Sperm should evolve to make female meiosis fair.

### 2014-05-19 writeup_and_notes/male_dependent_drive.tex
added more to description of repulsion phase in text

% \begin{figure}
% \includegraphics[width = 0.8 \textwidth]{Figures/trajectories_of_sperm_based_supressors.eps} 
% \caption{Trajectories of two sperm-based supressors of drive. Possibly
% merge this figure the 3 allele cartoon}  \label{Trajectories_of_supressors}
% \end{figure}

@@

We first assume that the modifier is tightly
        linked to the drive locus (effectively creating a third
        allele/haplotype at this locus) and arises on the drive-background. 
%We complement this single-locus approach %of an allele with pleiotropic effects in both sexes, 
%	with alternative models of two linked loci - one a female driver, 
%	and the other, a sperm allele which modifies the effect of drive upon fertilization. 
%In this model, a female meiotic driver with no sperm-dependency 
%	initially reaches drive-viability equilibrium (with two alleles
%	$A$ and $B$ are the ancestral non-driver and driver alleles). 
%Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
 %      linked to this balanced locus (effectively creating a third
 %       allele/haplotype at this locus). 
Tight linkage offers the best
        chance for a collaboration to evolve between a driver and
       an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}. 
We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
	in Model 5 the modifier arises tightly linked to the non-driving allele, 
	and in Model 6 it is unlinked to the driver.

%\yb{??In the SUPP we explore the dynamics of a sperm modifier of female drive unlinked to the female driver??}

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
%We first consider a sperm-acting modifier of
%	female meiotic drive tightly linked to the $B$ background (Model 4).
%We treat this non-recombining drive/modifier haplotype as a third allele, and we label this allele as $B^+$ when it enhances drive. 
%(i.e. the non-recombining haplotype of a
 %       driver allele and sperm-acting drive modifier)
%The $B^+$ allele 
%	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon})
%		generating a third `$B^{+}$' allele (i.e. the non-recombining haplotype of a
   %     driver allele and sperm-acting drive modifier). 
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
%	for a simple schematic of this drive model.  
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, %the sperm-acting drive facilitator (
	the novel $B^{+}$ haplotype
	%) 
	arises when the ancestral driver 
	%($B$) 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
%As $B^{+}$ facilitates $B$ these interactions frequently form 
%	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
\gc{Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. This new allele still
        experiences femaledrive, but does so at a lower rate (or not
        at all) when it
        is fertilized by itself.} This $B^{-}$ displaces the ancestral driver (Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.

>>>>

@@

We first assume that the modifier is tightly
        linked to the drive locus (effectively creating a third
        allele/haplotype at this locus) and arises on the drive-background. 
Tight linkage offers the best
        chance for a collaboration to evolve between a driver and
       an enhancer of the drive, as recombination breaks up drive haplotypes \citep{Thomson1974,Charlesworth1978,Haig1991}. 
Additionally, tight linkage between female driver and sperm modifier is consistent with the nature of well characterized drive systems which are often maintained as polymorphic inversions with numerous linked modifiers \cite{Burt2006}. 
We conclude by analyzing models with alternative linkage relationship between driver and drive modifier - 
	in Model 5 the modifier arises tightly linked to the non-driving allele, 
	and in Model 6 it is unlinked to the driver.

%%%sperm-acting facilitator
When the modifier of drive is tightly linked to the driver and
enhances drive (i.e. in coupling phase), 
	we label  this non-recombining drive/modifier haplotype as the $B^+$ allele.  
The $B^+$ allele acts in sperm to increase the effectiveness of drive for both
	the $B$ and  $B^{+}$ alleles in $AB$ and $AB^{+}$ heterozygotes (see Figure \ref{Eggsperm_3_allele_cartoon}, and Model 4 in the Supplementary Worksheet). 
Naively, $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, 
	the novel $B^{+}$ haplotype
	arises when the ancestral driver 
	is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population (Appendix Model 4, eqn. \eqref{coupling}, and Mathematical Supplement).

%%%sperm-acting drive suppressors
While  sperm enhancement of a female drive cannot displace a polymorphic female driver, sperm based drive suppressors can. 
%	that is sperm-acting alleles that act to restore  fairness of female meiosis, can. %If these sperm-acting drive suppressors arise on
\gc{Imagine a sperm-acting allele that restores fairness to female meiosis arises on
	the drive background, i.e. in repulsion phase, creating a
        third allele $B^{-}$, Model 4. \gc{This new allele still
        experiences female drive when fertilized by A or B sperm, but 
       is does not drive when fertilized by another $B^{-}$ so it
       avoid the excess formation of low fitness homozygotes.} This
     allows the $B^{-}$ to displace the ancestral driver (Figure
     \ref{Eggsperm_3_allele_cartoon}, Equation \ref{coupling}), 
	and often returns to a lower equilibrium frequency than the the $B$ allele, further decreasing the extent of drive in the population.  
If this sperm-acting drive suppressor arises on 
	the non-driving $A$ background (creating a third allele $A^{-}$, Model 5), 
	or is unlinked to the drive locus (Model 6), it readily invades a population segregating
	for the drive system (Figure \ref{Eggsperm_3_allele_cartoon} and eqns \ref{A-} and \ref{unlinked}).  
This allele lowers the frequency of the original driver (perhaps to zero),
	and spreads to fixation if it does not carry strong fitness costs
	(Figure \ref{Eggsperm_3_allele_cartoon} and \ref{Trajectories_of_supressors}A). 
This result is consistent with previous work showing that drive suppressors unlinked to, or in repulsion phase with drivers usually invade polymorphic drive systems \citep[e.g. ][]{Brandvain2012}.  
Therefore, all two-locus models of sperm influence on female drive suggest that
	sperm will evolve to oppose female meiotic drive.
